Atomify - a live LAMMPS visualizer

### 2017-04-07 ms.tex
Added placeholders for figures

\documentclass[aps,pre,twocolumn,letterpaper,floatfix]{revtex4}
\usepackage{graphicx} 
\usepackage{amsmath,amssymb,amsfonts} 
\usepackage{mathtools}
\usepackage{pdfpages}
\usepackage{afterpage}
\usepackage[hidelinks]{hyperref} 
\usepackage{epstopdf}
\begin{document}

@@

\section{Introduction}
%
Molecular dynamics (MD) is a common technique for simulating the motion of atoms
and molecules in a broad range of fields.
% TODO does MD also include larger particles?
% TODO which fields?
Over the years, increasingly sophisticated methods have been developed.
% TODO what methods?
With the large number of force fields, advanced techniques and GPU acceleration,
implementing a good molecular dynamics simulation is a challenging task.
There are many software packages for molecular dynamics modelling
available
\footnote{See \url{https://en.wikipedia.org/wiki/Comparison_of_software_for_molecular_mechanics_modeling}.}
Some packages, such as LAMMPS\cite{plimpton1995fast} and GROMACS\cite{Pronk2013}
have been developed for more than two decades and contain more than one million
lines of code each.

The work in this paper is focused on LAMMPS.
Although the software is well-documented and enables the user to perform
advanced simulations, it can be demanding to learn how to use it well.
A LAMMPS simulation is defined by a series of commands which are executed one by
one.
Physical quantities such as energy, temperature and stress can be measured by
adding \textit{computes}.
The time evolution of these quantities can in turn be saved to file for plotting
and further analysis.
Particle trajectories can be saved to a file and visualized by tools such as
VMD\cite{Humphrey1996Vmd} or OVITO\cite{Stukowski2009Visualization}.
The lack of an intuitive GUI and visualization \footnote{which are listed in the
non-features list on the web page:
\url{http://lammps.sandia.gov/non_features.html}} makes the workflow somewhat
complicated.

Atomify solves combines script editing, simulation,
visualization, and analysis in one single software application.
It uses LAMMPS as a library and synchronizes after each timestep,
copying the relevant information from LAMMPS to the graphics processing unit
(GPU) for visualization and data plotting.
The visualization is optimized for high performance by raytracing on billboards.
This allows visualization of millions of atoms with good framerate on desktop
computers.

@@

@@

@@

\subsection{Script editor}

Atomify provides a lightweight script editor with syntax highlighting and line
numbers.
% TODO add built-in help and autocompletion to Atomify and mention here
Multiple files can be open at the same time and persist across working sessions.
The script can easily be restarted by clicking the button in the GUI or using
the default shortcut (Ctrl+R in Windows and Linux or Cmd+R in macOS).
Any changes made to the script are immediately visualized in the simulation.

\subsection{Visualization}

@@

\subsection{Property coloring}

Measures of microscopic quantities can be visualized.
Any property of the individual atoms and bonds can in principle be used to color
the spheres and cylinders, respectively.
% TODO describe how this works in the GUI

@@

@@

\section{Case study}
% Rerun on large dataset

>>>>

\documentclass[aps,pre,twocolumn,letterpaper,floatfix]{revtex4}
\usepackage{graphicx} 
\usepackage{amsmath,amssymb,amsfonts} 
\usepackage{mathtools}
\usepackage{pdfpages}
\usepackage{afterpage}
\usepackage[hidelinks]{hyperref} 
\usepackage{epstopdf}
\usepackage{todo}
\begin{document}

@@

\section{Introduction}
%
Molecular dynamics (MD) is a common technique for simulating the motion of atoms
and molecules in a broad range of fields.
% TODO does MD also include larger particles?
% TODO which fields?
Over the years, increasingly sophisticated methods have been developed.
% TODO what methods?
With the large number of force fields, advanced techniques and GPU acceleration,
implementing a good molecular dynamics simulation is a challenging task.
There are many software packages for molecular dynamics modelling
available
\footnote{See \url{https://en.wikipedia.org/wiki/Comparison_of_software_for_molecular_mechanics_modeling}.}
Some packages, such as LAMMPS\cite{Plimpton1995Fast}, GROMACS\cite{Pronk2013}
and NAMD\cite{Phillips2005Scalable} have been developed for more than two decades 
and contain more than one million lines of code each.

The work in this paper is focused on LAMMPS.
Although the software is well-documented and enables the user to perform
advanced simulations, it can be demanding to learn how to use it well.
A LAMMPS simulation is performed by a series of commands which are executed one by
one.
Physical quantities such as energy, temperature and stress can be measured by
adding \textit{computes}.
The time evolution of these quantities can in turn be saved to file for plotting
and further analysis.
Particle trajectories can be saved to a file and visualized by tools such as
VMD\cite{Humphrey1996Vmd} or OVITO\cite{Stukowski2009Visualization}.
The lack of an intuitive GUI and visualization \footnote{which are listed in the
non-features list on the web page:
\url{http://lammps.sandia.gov/non_features.html}} makes the workflow somewhat
complicated.

Atomify solves combines script editing, simulation,
visualization, and analysis in one single software application.
It uses LAMMPS as a library and synchronizes after each timestep,
copying the relevant information from LAMMPS to the graphics processing unit
(GPU) for visualization and data plotting.
The visualization is optimized for high performance by raytracing on billboards.
This allows visualization of millions of atoms with good framerate on desktop
computers. Atomify is designed with simplicity as main goal.

TODO: Write about existing GUI software such as Materials Studio and MacroModel \\

@@

\begin{figure}
	\centering
	\includegraphics[width=0.5\textwidth]{final_billboard.png}
	\caption{Oversikt over GUI}
	\label{fig:gui}
\end{figure}

@@

\begin{figure}
	\centering
	\includegraphics[width=0.5\textwidth]{final_billboard.png}
	\caption{Workflow}
	\label{fig:gui}
\end{figure}

@@

\subsection{Script editor}
Atomify provides a lightweight script editor with syntax highlighting and line
numbers.
% TODO add built-in help and autocompletion to Atomify and mention here
Multiple files can be open at the same time and persist across working sessions.
The script can easily be restarted by clicking the button in the GUI or using
the default shortcut (Ctrl+R in Windows and Linux or Cmd+R in macOS).
Any changes made to the script are immediately visualized in the simulation.

\subsection{Visualization}
TODO: Color atoms based on property \\

\subsection{Dashboard}
TODO: Highlight groups/regions \\
TODO: Show/hide groups/regions \\
TODO: Rendering parameters \\
TODO: Atomify commands \\
TODO: Rendering parameters \\
TODO: Plot compute/variable/fix \\

\subsection{Live plotting}
TODO: Histogram over per atom quantities \\
TODO: Time evolution of quantity \\
TODO: 2d ave/chunk binning \\

@@

\begin{figure}
	\centering
	\includegraphics[width=0.5\textwidth]{final_billboard.png}
	\caption{Features}
	\label{fig:gui}
\end{figure}

@@

\begin{figure}
	\centering
	\includegraphics[width=0.5\textwidth]{final_billboard.png}
	\caption{Data pipeline (lammps -> atomify)}
	\label{fig:gui}
\end{figure}

@@

\section{Case study (Anders)}
% Rerun on large dataset

\begin{figure}
	\centering
	\includegraphics[width=0.5\textwidth]{final_billboard.png}
	\caption{Case study result}
	\label{fig:gui}
\end{figure}
